### 2020-06-19 article.tex
From .cpp to .c

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={KSPHPDDM and PCHPDDM: A Class of Adaptive Krylov and Overlapping Schwarz Methods in PETSc},
  pdfauthor={P. Jolivet and S. Zampini}
}
\fi

@@

% FundRef data to be entered by SIAM
%<funding-group specific-use="FundRef">
%<award-group>
%<funding-source>
%<named-content content-type="funder-name"> 
%</named-content> 
%<named-content content-type="funder-identifier"> 
%</named-content>
%</funding-source>
%<award-id> </award-id>
%</award-group>
%</funding-group>
\newcommand\col{:}
\newcommand{\pk}[1]{{\color{gray!90!black} #1}}
\usepackage{stmaryrd}
\usepackage{pgfplotstable}
\pgfplotsset{compat=newest}
\pgfplotsset{colormap={paraview}{rgb(0cm)=(0.278431,0.278431,0.858824) rgb(0.1428571429cm)=(0,0,0.360784) rgb(0.2857142857cm)=(0,1,1) rgb(0.4285714286cm)=(0,0.501961,0) rgb(0.5714285714cm)=(1,1,0) rgb(0.7142857143cm)=(1,0.380392,0) rgb(0.8571428571cm)=(0.419608,0,0) rgb(1cm)=(0.878431,0.301961,0.301961)}}
\usepackage{fancyvrb}
\usepackage{subfig}
\usepackage{upquote,textcomp}
\newcommand*{\fvtextcolor}[2]{\textcolor{#1}{#2}}
\definecolor{last}{RGB}{44,123,182}
\definecolor{blueaccent}{RGB}{215,25,28}
\definecolor{myyellow}{RGB}{102,102,17}
\definecolor{myred}{RGB}{197,65,0}
\definecolor{myblue}{RGB}{0,0,235}
\definecolor{myaqua}{RGB}{9,156,156}
\definecolor{mygreen}{RGB}{63,126,28}
\definecolor{paraviewred}{rgb}{0.278431,0.278431,0.858824}
\definecolor{paraviewblue}{rgb}{0.878431,0.301961,0.301961}
\definecolor{darkgray}{RGB}{253,174,97}
\definecolor{mediumgray}{RGB}{255,255,191}
\usepackage{tikz}
\usepackage[binary-units=true]{siunitx}
\usetikzlibrary{decorations.text,arrows.meta,calc,plotmarks,backgrounds,shadows,spy,fixedpointarithmetic,patterns,intersections,arrows.meta,shapes,decorations.text,decorations.pathmorphing,backgrounds,fit,positioning,shapes.symbols,chains,3d,calc,external}
\long\def\ifnodedefined#1#2#3{%
    \@ifundefined{pgf@sh@ns@#1}{#3}{#2}%
}
\makeatother
\usepackage{amsmath,etoolbox}
\AtBeginEnvironment{pmatrix}{\setlength{\arraycolsep}{2pt}}
\pgfmathsetmacro{\dof}{4}
\pgfmathsetmacro{\overlap}{1}
\begin{document}

@@

% REQUIRED
\begin{abstract}
  Contemporary applications in computational science and engineering often
    require the solution of linear systems which may be of different sizes,
    shapes, and structures. The goal of this paper is to explain how two
    libraries, PETSc and HPDDM, have been interfaced together in order to offer
    to end-users adaptive Krylov and overlapping Schwarz methods.  The
    flexibility of the implementation, which for example may deal with various
    matrix formats, are showcased and explained.  Diverse examples are
    displayed, ranging from fracture mechanics to linear stability analysis, and
    the sets of options are thoroughly documented, so that they may be readily
    tried on other applications.
\end{abstract}

@@

\section{Introduction}
Implementing and maintaining robust and scalable applications is highly relevant
in the field of computational science and engineering. Some tools exist to help
end-users achieve this goal. One the one hand, PETSc \cite{balay1997cient}, the
Portable and Extensible Toolkit for Scientific computation, is a
well-established software from the community. It may be used to efficiently
discretize partial differential equations, or solve algebraic linear or
nonlinear systems of equations. It comes with many built-in features such as
advanced preconditioning or tailored matrix formats. Also, it can be used to
compile and interact with third-party libraries, such as
\emph{hypre}~\cite{falgout2002hypre} for linear solvers, TetGen~\cite{si2013tet}
for mesh generation, and so on. These appealing features convinced computational
scientists to use PETSc as one of the discretization and/or algebraic backend in
many different higher-level projects, such as finite element or finite volume
packages. On the other hand, HPDDM is a much smaller project purely focused on
domain decomposition preconditioners~\cite{jolivet2013scalable}, a small
subclass of preconditioners available in PETSc as well, and advanced iterative
methods~\cite{jolivet2016block}. Such methods deal efficiently with linear
systems when multiple right-hand sides are available simultaneously, and are
then referred to as block iterative methods, or when there is a recurrence of
varying coefficient matrices and right-hand sides, and are then referred to as
recycling iterative methods.
% Htool ?
In this paper, the integration of both aspects of HPDDM, its advanced iterative
methods and its domain decomposition preconditioners, into PETSc is explained.
The design principles of the interface are showcased, alongside minimalist and
easy-to-run examples. The paper is divided in two main parts:
\cref{sec:KSPHPDDM} which introduces \pk{KSPHPDDM}, and \cref{sec:PCHPDDM} which
introduces \pk{PCHPDDM}. Eventually, concluding remarks may be found
\cref{sec:conclusion}. All PETSc keywords are highlighted in \pk{gray}, these
are public and documented at \url{https://www.mcs.anl.gov/petsc/petsc-master/docs}.
\section{KSPHPDDM}\label{sec:KSPHPDDM}
  \subsection{Related work}
Krylov subspace methods are widely used in numerical linear algebra for solving
systems of equations. Some of their strengths are memory efficiency and the fact
that they are mostly format agnostic. Indeed, they mainly rely on matrix--vector
operations such as multiplications or transpose multiplications. For
challenging high-dimensional systems, finding appropriate preconditioners is
mandatory to make these methods robust, but this goes beyond the scope of this
section. In PETSc, Krylov methods inherit from the \pk{KSP} base class. Some of the
most used types are \pk{KSPGMRES}, \pk{KSPCG}, or \pk{KSPBCGS}, which respectively implement
the generalized minimal residual method~\cite{saad1986gmres}~(GMRES), conjugate
gradient~\cite{hestenes1952methods}~(CG), or biconjugate gradient stabilized method~\cite{van1992bi}~(BCGS). Some of these methods are also
available in other linear algebra backends such as~\emph{hypre} or DUNE~\cite{bastian2010generic}.

@@

Recycling put aside, another important aspect of Krylov methods is their ability
to deal with multiple right-hand sides simultaneously. Block Krylov methods~\cite{gutknecht2006block} are
designed for solving linear systems $A X = B$, where
$X$ and $B$ are tall-and-skinny matrices with $k \geq 1$ columns. These methods are
less frequently implemented in general purpose libraries because they require
somehow more involved kernels such as matrix--matrix multiplication, instead of
matrix--vector. Still, systems with multiple right-hand sides are ubiquitous,
e.g., in tomography~\cite{tournier2016micro} or data
analytics~\cite{KALANTZIS2018136}. Of course, it is possible to cast a system
with multiple right-hand sides into a sequence of systems with the same
coefficient matrix but changing right-hand sides $A x_i = b_i, i =
1,2,\ldots,k$. However, block Krylov methods, at the cost of higher arithmetic
intensities, generate larger subspaces and typically converge in fewer iterates.
There is none implemented in PETSc. It is nevertheless possible to solve
systems with multiple right-hand sides using a specialized type of matrices,
but with a very limited choice of preconditioners, cf.~\cref{sec:matkaij}.
In the context of optimization, it is also
frequent to solve a sequence of linear systems $A_i X_i = B_i, i = 1, 2,
\ldots$

@@

Trilinos interoperate is possible, there is currently no \pk{KSP} interface to Belos.
This limits its composability inside complete PETSc applications. Furthermore,

@@

@@

\pk{KSPHPDDM}) and its Fortran equivalent.

@@

Unfortunately, PETSc does not have a general routine for applying $P_B$ to a

@@

N    baij-1-T    baij-1-F   sbaij-1-T   sbaij-1-F  aij-1-T     aij-1-F    baij-3-T    baij-3-F   sbaij-3-T   sbaij-3-F  aij-3-T     aij-3-F    baij-10-T   baij-10-F  sbaij-10-T  sbaij-10-F aij-10-T    aij-10-F
1    1.8027e-01  1.4696e+08 6.1586e-02  1.4696e+08 8.0527e-02  1.4696e+08 7.1397e-01  1.3535e+09 5.1037e-01  1.3638e+09 1.0686e+00  1.3535e+09 6.1909e+00  1.5159e+10 4.6958e+00  1.5206e+10 1.6114e+01  1.5159e+10
2    3.5192e-01  2.9392e+08 1.6650e-01  3.2482e+08 1.8615e-01  3.0422e+08 1.0437e+00  2.7071e+09 9.3575e-01  2.9234e+09 1.4753e+00  2.7380e+09 1.4970e+01  3.0319e+10 1.5530e+01  3.2482e+10 1.6267e+01  3.0422e+10
8    6.1599e-01  1.1757e+09 7.2168e-01  1.2993e+09 3.1676e-01  1.2169e+09 2.6101e+00  1.0828e+10 3.0701e+00  1.1694e+10 2.0870e+00  1.0952e+10 2.7886e+01  1.2128e+11 3.4780e+01  1.2993e+11 2.2569e+01  1.2169e+11
16   1.9594e+00  2.3513e+09 2.6186e+00  2.5986e+09 6.2770e-01  2.4337e+09 7.5274e+00  2.1656e+10 9.8057e+00  2.3387e+10 4.1388e+00  2.1904e+10 3.6085e+01  2.4255e+11 3.6290e+01  2.5986e+11 4.5153e+01  2.4337e+11
32   4.3453e+00  4.7026e+09 5.4669e+00  5.1972e+09 1.2538e+00  4.8675e+09 1.6166e+01  4.3313e+10 2.0638e+01  4.6775e+10 8.1531e+00  4.3807e+10 5.2586e+01  4.8510e+11 5.2132e+01  5.1972e+11 1.0705e+02  4.8675e+11
64   9.6799e+00  9.4053e+09 1.1616e+01  1.0394e+10 2.5713e+00  9.7350e+09 3.4513e+01  8.6626e+10 4.2465e+01  9.3549e+10 1.6444e+01  8.7615e+10 8.3267e+01  9.7020e+11 8.6683e+01  1.0394e+12 2.2792e+02  9.7350e+11

@@

The Poisson equation on a cube is discretized with $\mathbb{P}_1$ finite elements in double precision. This yields a matrix $A$ of dimension one million. Though the numerical values of $A$ are of no interest here, the sparsity pattern is rather common and frequently encountered, e.g., when discretizing partial differential equations. Then, a random dense matrix $T$ of dimension $b$ is generated, for $b \in \{ 1,3,10 \}$, and the following matrix is evaluated $\mathbb{A} = A \otimes T$. The matrix $\mathbb{A}$ is then assembled into all three formats described above. For block matrices, a block size of $b$ is used. The performance of the \pk{MatProductNumeric} operation is evaluated for tall-and-skinny dense matrices with a varying number of columns $N \in \{1,2,8,16,32,64\}$. For $N=1$, the \pk{MatMult} operation for matrix--vector multiplication is used instead. All results are gathered in~\cref{fig:perfAX}. They can be reproduced using the following mini-app: \url{http://jolivet.perso.enseeiht.fr/MatProduct.cpp} and any input \pk{MatAIJ} stored in binary format, here using the default name binaryoutput.

@@

&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpicxx] MatProduct.cpp -O3 -std=c++11 -I&fvtextcolor[myyellow][${PETSC_DIR}]/&fvtextcolor[myyellow][${PETSC_ARCH}]/include                     &fvtextcolor[myred][\]

@@

  -bs 1,3,10 -N 1,2,8,16,32,64 -type aij,baij,sbaij

@@

@@

        enlarge x limits=0.1, height=4cm,
        width=0.98\textwidth,legend columns=4,
        legend style={at={(0.5,1.65)}, legend cell align=left,
        anchor=north},

@@

        xtick = {1,2,4,8,16,32},
ytick = {0.2,1,2.5,5},ylabel style={yshift=6pt},xtick pos=left,

@@

@@

          cycle list={blueaccent, darkgray, mediumgray},%, lightgray,last},

@@

>>>>

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={KSPHPDDM and PCHPDDM: Two Classes of Adaptive Krylov and Overlapping Schwarz Methods in PETSc},
  pdfauthor={P. Jolivet and S. Zampini}
}
\fi

@@

% FundRef data to be entered by SIAM
%<funding-group specific-use="FundRef">
%<award-group>
%<funding-source>
%<named-content content-type="funder-name"> 
%</named-content> 
%<named-content content-type="funder-identifier"> 
%</named-content>
%</funding-source>
%<award-id> </award-id>
%</award-group>
%</funding-group>
\newcommand\col{:}
\newcommand{\pk}[1]{{\color{gray!90!black} #1}}
\usepackage{stmaryrd}
\usepackage{pgfplotstable}
\pgfplotsset{compat=newest}
\pgfplotsset{colormap={paraview}{rgb(0cm)=(0.278431,0.278431,0.858824) rgb(0.1428571429cm)=(0,0,0.360784) rgb(0.2857142857cm)=(0,1,1) rgb(0.4285714286cm)=(0,0.501961,0) rgb(0.5714285714cm)=(1,1,0) rgb(0.7142857143cm)=(1,0.380392,0) rgb(0.8571428571cm)=(0.419608,0,0) rgb(1cm)=(0.878431,0.301961,0.301961)}}
\usepackage{fancyvrb}
\usepackage{subfig}
\usepackage{upquote,textcomp}
\newcommand*{\fvtextcolor}[2]{\textcolor{#1}{#2}}
\definecolor{last}{RGB}{44,123,182}
\definecolor{blueaccent}{RGB}{215,25,28}
\definecolor{myyellow}{RGB}{102,102,17}
\definecolor{myred}{RGB}{197,65,0}
\definecolor{myblue}{RGB}{0,0,235}
\definecolor{myaqua}{RGB}{9,156,156}
\definecolor{mygreen}{RGB}{63,126,28}
\definecolor{paraviewred}{rgb}{0.278431,0.278431,0.858824}
\definecolor{paraviewblue}{rgb}{0.878431,0.301961,0.301961}
\definecolor{darkgray}{RGB}{253,174,97}
\definecolor{mediumgray}{RGB}{255,255,191}
\usepackage[outline]{contour}
\usepackage{tikz}
\usepackage[binary-units=true]{siunitx}
\usetikzlibrary{decorations.text,arrows.meta,calc,plotmarks,shadows,spy,fixedpointarithmetic,patterns,intersections,arrows.meta,shapes,decorations.text,decorations.pathmorphing,backgrounds,fit,positioning,shapes.symbols,chains,3d,calc,external}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\long\def\ifnodedefined#1#2#3{%
    \@ifundefined{pgf@sh@ns@#1}{#3}{#2}%
}
\makeatother
\usepackage{amsmath,etoolbox}
\AtBeginEnvironment{pmatrix}{\setlength{\arraycolsep}{2pt}}
\pgfmathsetmacro{\dof}{4}
\pgfmathsetmacro{\overlap}{1}
\begin{document}

@@

% REQUIRED
\begin{abstract}
  Contemporary applications in computational science and engineering often
    require the solution of linear systems which may be of different sizes,
    shapes, and structures. The goal of this paper is to explain how two
    libraries, PETSc and HPDDM, have been interfaced together in order to offer
    to end-users adaptive Krylov and overlapping Schwarz methods.  The
    flexibility of the implementation %, which for example may deal with various matrix formats, 
    are showcased and explained.  Diverse examples are
    displayed, ranging from fracture mechanics to linear stability analysis, and
    the sets of options are thoroughly documented, so that they may be readily
    tried on other applications and reproduce results.
\end{abstract}

@@

\section{Introduction}
Implementing and maintaining robust and scalable applications is highly relevant
in the field of computational science and engineering. Some tools exist to help
end-users achieve this goal. One the one hand, PETSc \cite{balay1997cient}, the
Portable and Extensible Toolkit for Scientific computation, is a
well-established software from the community. It may be used to efficiently
discretize partial differential equations, or solve algebraic linear or
nonlinear systems of equations. It comes with many built-in features such as
advanced preconditioning or tailored matrix formats. Also, it can % be used to compile and 
interact with third-party libraries, such as
\emph{hypre}~\cite{falgout2002hypre} for linear solvers, TetGen~\cite{si2013tet}
for mesh generation, and so on. These appealing features convinced computational
scientists to use PETSc as one of the discretization and/or algebraic backend in
many different higher-level projects, such as finite element or finite volume
packages. On the other hand, HPDDM is a much smaller project purely focused on
domain decomposition preconditioners~\cite{jolivet2013scalable} % , a small subclass of preconditioners available in PETSc as well, 
and advanced iterative
methods~\cite{jolivet2016block}. Such methods deal efficiently with linear
systems when multiple right-hand sides are available simultaneously, and are
then referred to as block iterative methods, or when there is a recurrence of
varying coefficient matrices and right-hand sides, and are then referred to as
recycling iterative methods.
% Htool ?
In this paper, the integration of both aspects of HPDDM, its advanced iterative
methods and its domain decomposition preconditioners, into PETSc is explained.
The design principles of the interface are showcased, with % alongside
minimalist and easy-to-run examples. The paper is divided in two main parts:
\cref{sec:KSPHPDDM} which introduces \pk{KSPHPDDM}, and \cref{sec:PCHPDDM} which
introduces \pk{PCHPDDM}. Eventually, concluding remarks may be found
\cref{sec:conclusion}. All PETSc keywords are highlighted in \pk{gray}, these
are public and documented at \url{https://www.mcs.anl.gov/petsc/petsc-master/docs}.
\section{KSPHPDDM}\label{sec:KSPHPDDM}
  \subsection{Related work}
Krylov subspace methods are widely used in numerical linear algebra for solving
systems of equations. Some of their strengths are memory efficiency and the fact
that they are mostly format agnostic. Indeed, they mainly rely on matrix--vector
operations such as multiplications or transpose multiplications. For
challenging high-dimensional systems, finding appropriate preconditioners is
mandatory to make these methods robust. %, but this goes beyond the scope of this section.
In PETSc, Krylov methods belong to the \pk{KSP} base class. Some of the
most used types are \pk{KSPGMRES}, \pk{KSPCG}, or \pk{KSPBCGS}, which respectively implement
the generalized minimal residual method~\cite{saad1986gmres}~(GMRES), conjugate
gradient~\cite{hestenes1952methods}~(CG), or biconjugate gradient stabilized method~\cite{van1992bi}~(BCGS). % Some of these methods are also
% available in other linear algebra backends such as~\emph{hypre} or DUNE~\cite{bastian2010generic}.

@@

% Recycling put aside, 
Another important aspect of Krylov methods is their ability
to deal with multiple right-hand sides simultaneously. Block Krylov methods~\cite{gutknecht2006block} are
designed for solving linear systems $A X = B$, where
$X$ and $B$ are tall-and-skinny matrices with $k \geq 1$ columns. These methods are
less frequently implemented in general purpose libraries because they require
somehow more involved kernels such as matrix--matrix multiplication, instead of
matrix--vector. Still, systems with multiple right-hand sides are ubiquitous,
e.g., in tomography~\cite{tournier2016micro} or data
analytics~\cite{KALANTZIS2018136}. Of course, it is possible to cast a system
with multiple right-hand sides into a sequence of systems with the same
coefficient matrix but changing right-hand sides $A x_i = b_i, i =
1,2,\ldots,k$. However, block Krylov methods, at the cost of higher arithmetic
intensities, generate larger subspaces and typically converge in fewer iterates.
There is none implemented in PETSc. It is nevertheless possible to solve
systems with multiple right-hand sides using a specialized type of matrices,
but with a very limited choice of preconditioners, cf.~\cref{sec:matkaij}.
In the context of optimization, it is also
frequent to solve a sequence of linear systems $A_i X_i = B_i, i = 1, 2,
\ldots$

@@

Trilinos interoperate is possible, there is currently no \pk{KSP} interface to Belos. % This limits its composability inside complete PETSc applications.
Furthermore,

@@

registered since PETSc version 3.12 when configuring PETSc with the extra

@@

\pk{KSPHPDDM}).% and its Fortran equivalent.

@@

Currently, PETSc does not have a general routine for applying $P_B$ to a

@@

N    baij-1-T    baij-1-F   sbaij-1-T   sbaij-1-F  aij-1-T     aij-1-F    baij-3-T    baij-3-F   sbaij-3-T   sbaij-3-F  aij-3-T     aij-3-F    baij-10-T   baij-10-F  sbaij-10-T  sbaij-10-F aij-10-T    aij-10-F     cusparse-1-T    cusparse-1-F    cusparse-3-T    cusparse-3-F    cusparse-10-T   cusparse-10-F    aijmkl-seq-1-T      aijmkl-seq-3-T      aijmkl-seq-10-T     baijmkl-seq-1-T      baijmkl-seq-3-T      baijmkl-seq-10-T   sbaijmkl-seq-1-T      sbaijmkl-seq-3-T      sbaijmkl-seq-10-T         aijmkl-1-T      aijmkl-3-T      aijmkl-10-T     baijmkl-1-T      baijmkl-3-T      baijmkl-10-T   sbaijmkl-1-T      sbaijmkl-3-T      sbaijmkl-10-T     
1      2.0126e-01   1.4696e+08 8.0098e-02   1.4696e+08   9.3388e-02   1.4696e+08     8.9584e-01   1.3535e+09    6.2428e-01   1.3638e+09     1.2072e+00   1.3535e+09   4.0734e+00   5.4450e+09    2.2476e+00   5.4708e+09    5.0428e+00   5.4450e+09   4.4800e-02   1.5211e+08   3.7933e-01   1.3690e+09   1.5043e+00   5.4759e+09   6.6857e-02  7.4649e-01   3.1128e+00   4.7672e-01   6.2903e-01  2.2271e+00   4.7675e-01   6.2990e-01    2.2273e+00    7.8399e-03    8.6487e-02    3.5536e-01   5.2814e-02 6.6225e-02   2.4584e-01   4.5115e-02   6.6404e-02   2.4489e-01
2      3.2635e-01   2.9392e+08 2.4204e-01   3.2482e+08   1.1522e-01   3.0422e+08     1.2835e+00   2.7071e+09    1.0697e+00   2.9234e+09     9.9480e-01   2.7380e+09   1.4504e+01   1.0890e+10    1.5265e+01   1.1694e+10    3.7799e+00   1.0952e+10   5.2164e-03   3.0422e+08   3.4056e-02   2.7380e+09   7.7659e-02   1.0952e+10   1.1703e-01  9.3872e-01   3.5976e+00   4.7905e-01   7.3497e-01  3.3331e+00   4.7886e-01   7.3415e-01    3.3297e+00    1.2450e-02    1.0124e-01    3.9873e-01   4.7633e-02 7.4107e-02   2.8513e-01   4.7733e-02   7.4693e-02   2.8507e-01
8      7.9573e-01   1.1757e+09 1.0070e+00   1.2993e+09   3.3729e-01   1.2169e+09     3.3663e+00   1.0828e+10    3.9830e+00   1.1694e+10     2.4656e+00   1.0952e+10   2.2209e+01   4.3560e+10    2.2545e+01   4.6775e+10    9.2040e+00   4.3807e+10   1.3652e-02   1.2169e+09   8.2139e-02   1.0952e+10   1.8984e-01   4.3807e+10   4.2722e-01  2.8760e+00   1.0913e+01   1.3256e+00   1.7457e+00  7.9644e+00   1.3210e+00   1.7410e+00    7.9679e+00    4.2997e-02    2.6899e-01    1.6814e+00   1.3183e-01 1.7832e-01   8.8593e-01   1.3082e-01   1.7947e-01   8.8420e-01
16     2.1378e+00   2.3513e+09 3.0837e+00   2.5986e+09   6.7295e-01   2.4337e+09     9.1522e+00   2.1656e+10    1.1762e+01   2.3387e+10     4.9326e+00   2.1904e+10   3.1530e+01   8.7120e+10    2.9334e+01   9.3549e+10    1.8462e+01   8.7615e+10   2.4907e-02   2.4337e+09   1.3832e-01   2.1904e+10   3.3609e-01   8.7615e+10   8.5514e-01  5.7493e+00   2.1980e+01   2.6391e+00   3.4887e+00  1.5947e+01   2.6455e+00   3.4893e+00    1.5928e+01    8.4255e-02    5.4284e-01    3.3861e+00   2.6162e-01 3.5777e-01   1.7955e+00   2.6238e-01   3.5880e-01   1.7923e+00
32     4.2781e+00   4.7026e+09 6.5426e+00   5.1972e+09   1.3447e+00   4.8675e+09     1.8545e+01   4.3313e+10    2.4015e+01   4.6775e+10     9.8656e+00   4.3807e+10   4.4678e+01   1.7424e+11    4.1021e+01   1.8710e+11    3.6822e+01   1.7523e+11   4.7533e-02   4.8675e+09   2.6294e-01   4.3807e+10   6.3836e-01   1.7523e+11   1.7075e+00  1.1487e+01   4.3644e+01   5.2831e+00   6.9603e+00  3.1855e+01   5.2907e+00   6.9739e+00    3.1852e+01    1.6480e-01    1.0807e+00    6.7653e+00   5.2012e-01 7.1464e-01   3.5401e+00   5.2109e-01   7.1542e-01   3.5382e+00
64     9.9341e+00   9.4053e+09 1.3822e+01   1.0394e+10   2.6927e+00   9.7350e+09     3.5490e+01   8.6626e+10    4.6966e+01   9.3549e+10     1.9724e+01   8.7615e+10   6.7668e+01   3.4848e+11  6.3229e+01   3.7420e+11  7.4386e+01   3.5046e+11   8.3249e-02   9.7350e+09   5.0956e-01   8.7615e+10   1.2334e+00   3.5046e+11   3.4264e+00  2.3052e+01   8.7358e+01   1.0573e+01   1.3960e+01  6.3695e+01   1.0581e+01   1.3961e+01    6.3949e+01    3.2990e-01    2.1734e+00    1.3526e+01   1.0453e+00 1.4281e+00   7.0763e+00   1.0497e+00   1.4326e+00   7.1714e+00

@@

The Poisson equation on a cube is discretized with $\mathbb{P}_1$ finite elements in double precision. This yields a matrix $A$ of dimension one million. Though the numerical values of $A$ are of no interest here, the sparsity pattern is rather common and frequently encountered, e.g., when discretizing partial differential equations. Then, a random dense matrix $T$ of dimension $b$ is generated, for $b \in \{ 1,3,6 \}$, and the following matrix is evaluated $\mathbb{A} = A \otimes T$. The matrix $\mathbb{A}$ is then assembled into all three formats described above. For block matrices, a block size of $b$ is used. The performance of the \pk{MatProductNumeric} operation is evaluated for tall-and-skinny dense matrices with a varying number of columns $N \in \{1,2,8,16,32,64\}$. For $N=1$, the \pk{MatMult} operation for matrix--vector multiplication is used instead. All results are gathered in~\cref{fig:perfAX}. They can be reproduced using the following mini-app: \url{http://jolivet.perso.enseeiht.fr/MatProduct.cpp} and any input \pk{MatAIJ} stored in binary format, here using the default name binaryoutput.

@@

&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpicc] MatProduct.c -O3 -I&fvtextcolor[myyellow][${PETSC_DIR}]/&fvtextcolor[myyellow][${PETSC_ARCH}]/include                        &fvtextcolor[myred][\]

@@

  -bs 1,3,6 -N 1,2,8,16,32,64 -type aij,aijmkl,baij,baijmkl,sbaij,sbaijmkl

@@

  \pgfmathparse{1.4696e+08/9.3388e-02/1000000000.0}\edef\storeresultI{\pgfmathresult}
  \pgfmathparse{1.3535e+09/8.9584e-01/1000000000.0}\edef\storeresultII{\pgfmathresult}
  \pgfmathparse{5.4450e+09/4.0734e+00/1000000000.0}\edef\storeresultIII{\pgfmathresult}
  % \hspace*{-1.5cm}

@@

        enlarge x limits=0.101,height=4cm,
        width=0.8\textwidth,
        ylabel={},
        legend columns=3, legend style={legend cell align=left,at={(0.5,1.03)},anchor=south},
        xtick = {\empty},
        xmajorgrids=true,
        extra x ticks = {1,2,4,8,16,32},
        extra x tick labels = {\empty},
        extra tick style={
            grid style={dotted,black}
        xtick pos=left,name=plot1,
        xticklabels={\empty},xmajorticks=false,
        ytick = {0.5,1,1.5,2},
          cycle list={blueaccent, darkgray, mediumgray},%,last},
          clip=false,point meta=exp(x),
          \addplot+[nodes near coords style={xshift=0.5cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultI~GFLOP/s\fi}}},
draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-1-F}/\thisrow{aij-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aij1}
          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-1-F}/\thisrow{baij-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baij1}
          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-1-F}/\thisrow{sbaij-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaij1}
           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-1-F}/\thisrow{aijmkl-seq-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklaijseq1}
           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-1-F}/\thisrow{baijmkl-seq-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklbaijseq1}
           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-1-F}/\thisrow{sbaijmkl-seq-1-T}/(\storeresultI*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklsbaijseq1}
           \addlegendentry{\pk{MatAIJMKL}}
           \addlegendentry{\pk{MatBAIJMKL}}
           \addlegendentry{MatSBAIJMKL}
\node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=1$};
        enlarge x limits=0.101,height=4cm,
        width=0.8\textwidth,
        % legend columns=1, legend style={legend cell align=left,at={(1.02,0.5)},anchor=west},
        xtick = {\empty},
        xmajorgrids=true,
        extra x ticks = {1,2,4,8,16,32},
        extra x tick labels = {\empty},
        extra tick style={
            grid style={dotted,black}
        xtick pos=left,name=plot2,at={($(plot1.south)-(0,0.2cm)$)},anchor=north,
        xticklabels={\empty},xmajorticks=false,
        ytick = {1,2,3,4},
          cycle list={blueaccent, darkgray, mediumgray},%,last},
          clip=false,point meta=exp(x)
          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{aij-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:aij3}
          \addplot+[nodes near coords style={xshift=0.5cm,yshift=-0.1cm,above}, nodes near coords*={\rotatebox{45}{\tiny\color{black}{\pgfmathfloattofixed{\pgfplotspointmeta}\pgfmathparse{int(round(\pgfmathresult))}\ifnum\pgfmathresult=1\pgfmathprintnumber\storeresultII~GFLOP/s\fi}}},
              draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-3-F}/\thisrow{baij-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:baij3}
          \addplot+[draw=black,fill, area legend,postaction={pattern=crosshatch dots}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-3-F}/\thisrow{sbaij-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:sbaij3}
           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{aij-3-F}/\thisrow{aijmkl-seq-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklaijseq3}
           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{baij-3-F}/\thisrow{baijmkl-seq-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklbaijseq3}
           \addplot+[draw=black,fill, area legend,postaction={pattern=bricks}] table [x expr=\thisrow{N}>7?\thisrow{N}/2:\thisrow{N},y expr=\thisrow{sbaij-3-F}/\thisrow{sbaijmkl-seq-3-T}/(\storeresultII*1000000000.0)]{\loadedMatProduct};\label{pgfplots:mklsbaijseq3}
           % \addlegendentry{\pk{MatAIJ}}
           % \addlegendentry{\pk{MatBAIJ}}
           % \addlegendentry{\pk{MatSBAIJ}}
           % \addlegendentry{\pk{MatAIJMKL}}
           % \addlegendentry{\pk{MatBAIJMKL}}
           % \addlegendentry{MatSBAIJMKL}
\node[rotate=0,fill=white,inner sep=0pt,anchor=west] at (rel axis cs:1.02, 0.5) {$b=3$};
        enlarge x limits=0.101,height=4cm,
        width=0.8\textwidth,

@@

        ylabel={},
        xtick = {\empty},
        xmajorgrids=true,
        extra x ticks = {1,2,4,8,16,32},
        extra x tick labels = {1,2,8,16,32,64},
        extra tick style={
            grid style={dotted,black}
        xtick pos=left,name=plot3,at={($(plot2.south)-(0,0.2cm)$)},anchor=north,

@@

        ytick = {1,2,3,4},

@@

          cycle list={blueaccent, darkgray, mediumgray},%,last},
          clip=false, point meta = exp(x)

@@

      \contourlength{.6pt}
